A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-01-10 doc/large_strain.tex
a bit more on LIKWID

{\color{red}
Figure \ref{fig:roofline} shows the Roofline performance model.
As expected, the matrix-based approach is memory bound and (on average) achieves the performance of $8.5\,\rm{GFlop/s}$.
Matrix-free implementation of fine-strain elastic tangent operators (on average) leads to a much better performance of
$22.6\,\rm{GFlop/s}$, as well as about an order of magnitude higher computational intensity, but still stays in the memory bound regime.
The achieved performance is about 13\% of the peak arithmetic performance, which is considerably lower than the
70\% reported for the Laplace operation in \cite{kronbichler12}.
Further detailed studies are necessary to explain this behavior.
}

{\color{red}
To summarize, we can attribute the speedup of the matrix-free finite-strain tangent operators for higher order bases to two factors.
First, is the higher algorithmic intensity as compared to the matrix-based implementation, which is clearly a memory
bound, see Figure XXXX.
Second, is the reduction in the number of floating point operations per DoF thanks to the sum factorization technique, as discussed in
Section \ref{sec:mf} and in the reference \cite{kronbichler12}.
}

@@

In this contribution, we proposed and numerically investigated {\color{red}on the node level three different} matrix-free implementations of tangent operators for finite-strain elasticity with heterogeneous materials.
{\color{red}
To the best of our knowledge, this is the first work that applies matrix-free sum factorization techniques to such materials.}
{\color{red}Among the three algorithms, }
the implementation that caches the material part of the fourth-order spatial tangent stiffness tensor together with the second-order Kirchhoff stress was shown to be faster than the matrix-based approach for quadratic elements in 3D and cubic {\color{red}elements in 2D}.
{\color{red}
Given that in this case we did not utilize the specific form of the material part of the stiffness tensor, we conclude}
that the matrix-free implementation of the tangent operator can be applied to any constitutive model which operates with the material part of the fourth-order spatial tangent stiffness tensor on the quadrature point level.
{\color{red}
Roofline model indicated the matrix-free finite-strain elasticity tangent operators have an order magnitude higher algorithmic intensity. Further studies are required to identify how to further improve the intensity and reach the compute bound case.}

>>>>

{\color{red}
Figure \ref{fig:roofline} shows the Roofline performance model.
As expected, the matrix-based approach is memory bound and (on average) achieves the performance of $8.5\,\rm{GFlop/s}$.
Matrix-free implementation of fine-strain elastic tangent operators (on average) leads to a much better performance of
$22.6\,\rm{GFlop/s}$, as well as about an order of magnitude higher computational intensity, but still stays in the memory bound regime.
The achieved performance is about 13\% of the peak arithmetic performance, which is considerably lower than the
70\% reported for the Laplace operation in \cite{kronbichler12}.
When comparing the data for ``tensor4'' caching strategy (Algorithm \ref{alg:mf_tensor4}) to that from the ``tensor2'' (Algorithm \ref{alg:mf_tensor2}),
we observe very marginal increase in computational intensity and almost unchanged performance.
This indicates the origin for suboptimal results of matrix-free approach is most likely not related to the algorithms studied here.
Further detailed studies are necessary to explain this behavior.
}

{\color{red}
To summarize, we can attribute the speedup of the matrix-free finite-strain tangent operators for higher order bases to two factors.
First, is the higher algorithmic intensity as compared to the matrix-based implementation, which is clearly memory
bound, see Figure \ref{fig:roofline}.
Second, is the reduction in the number of floating point operations per DoF thanks to the sum factorization technique, as discussed in
Section \ref{sec:mf} and \cite[Section 2.4]{kronbichler12}.
}

@@

In this contribution, we proposed and numerically investigated {\color{red}on the node level three different} matrix-free implementations of tangent operators for finite-strain elasticity with heterogeneous materials.
{\color{red}
To the best of our knowledge, this is the first work that applies matrix-free sum factorization techniques to such materials.}
{\color{red}Among the three algorithms, }
the implementation that caches the material part of the fourth-order spatial tangent stiffness tensor together with the second-order Kirchhoff stress was shown to be faster than the matrix-based approach for quadratic elements in 3D and cubic {\color{red}elements in 2D}.
{\color{red}
Given that in this case we did not utilize the specific form of the material part of the stiffness tensor, we conclude}
that the matrix-free implementation of the tangent operator can be applied to any constitutive model which operates with the material part of the fourth-order spatial tangent stiffness tensor on the quadrature point level.
{\color{red}
Roofline model indicated the matrix-free finite-strain elasticity tangent operators have an order magnitude higher algorithmic intensity. Further studies are required to identify means to improve the computational intensity and performance of the matrix-free implementation.}
